\documentclass[a4paper]{article}
%\documentclass{report} % Can be article, report, book, anything you want!

\usepackage{hyperref}
\usepackage{vub} % this will automatically set to A4 paper. We're not in the US.
% if, for some obscure reason, you need another format, please file an issue (see further).

% The `vub` package has many options.
% Please read the documentation at https://gitlab.com/rubdos/texlive-vub if you want to learn about them.

% Feel free to mix and match with other packages.

\setlength{\parskip}{1em}
\title{Pharo project}
\subtitle{Metaprogramming and Reflection}
\faculty{Sciences and Bio-Engineering Sciences} % Note: without the word "Faculty"!
\author{Gérard Lichtert\\
        0557513\\
    \href{mailto:gerard.lichtert@vub.be}{gerard.lichtert@vub.be}
}
\date{\today}

\begin{document}
\maketitle
\tableofcontents
\newpage
\raggedright

\section{Overview}
This project consists of three parts.
\begin{itemize}
	\item Protected Visibility Modifier
	\item Object-Centric Debugging
	\item Multiple exception handlers
\end{itemize}
Each part has it's own .st file located in the /src directory.
Do note, that I haven't overriden Pharo's default behavior. For instance, if you
would like to use the protected Visbility Modifier, you will have to add the
plugins to the compiler prior to testing it out. To make testing easier, there
are already tests present that should cover the minimum implementation
requirements, and test that the different component (such as plugins) indeed
work as described.
\par
Some tests for the Object-Centric Debugging part of the assignment will always
pass due to it requireing for the debugger to open. In this case you should
manually check if the debugger indeed opens as intended or not when it
shouldn't. There are comments to inform the user when it should or shouldn't.
\par
You should be able to file in the .st files by dragging them in a Pharo 12.0
image, selecting ``changelist browser", followed by ``select all" and then
``file in" selected. Note that if a class normally used in Pharo was moved due to
an addition or modification of a method, a warning may pop up.
\section{Protected Visibility Modifier}
\subsection{Requirement 1}
The first requirement was to extend the compiler to support compilation of
$<$protected$>$ methods, using double registration of public methods and selector
mangling of self sends.\par
To achieve this, first we created a plugin for the compiler that mangles the
selector of methods that have the $<$protected$>$ pragma.
We also need double registration, to achieve this, I traced the method install to the point
where methods are installed in a class. This was in ClassDescription, which is
moved to the package since the method to install methods was modified.
\par
A method is installed with, among
others, a selector and a compiledMethod, to which we can ask if it has a
$<$protected$>$ pragma. If it does, we know it will already have a mangled
selector. In this case it should be installed normally.
However, if it does not have the $<$protected$>$ pragma, we know it is a public
method and therefore has to be installed twice, once with the normal selector
and once with the mangled selector. Since a selector is required to be passed
along prior to installation, we install it once normally and install it once but
with the mangled selector. Essentially we add another message send but with the
mangled selector.
\par
To mangle self and super send, we create another plugin, just like we already
did for the selector manging of protected methods. However, this time we will
mangle the messages sent to self and super receivers.
\subsection{Requirement 2}
Another requirement is that the modifier should be backwards compatible, meaning
that sometimes objects will already exist without the mangled selector, or
taking the $<$protected$>$ modifier into account. To overcome this, we need to
be extra careful when mangling self and super sends. More concretely, the plugin
will now check if the class (or superclass, depending on the receiver) has the
mangled method. If it does, it will mangle it, otherwise it will leave it as is.
\section{Object-Centric Debugging}
\subsection{Requirement}
For this part I was tasked with
extending Pharo to support object-centric debugging based on the approach of Costiou et al.
My implementation supports
\begin{itemize}
	\item breaking on every message send
	\item breaking on every message send in the passed method
	\item breaking on every message send in the passed method where the message
	      meets a certain condition.
	\item clearing all breakpoints
\end{itemize}
To implement this, I used anonymousSubclass to decouple the object from the
class (so it only breaks on that specific object). Furthermore I used MetaLinks
to create the breakpoints where needed. To give an example; since it's the most
specific method. When onMessageSendIn: \# greet when: [ :send $\mid$ send selector =
\#, ]
is sent to the objectBreaker. The objectBreaker will compile the method from the
original class in the anonymousSubclass, go through the ast of the method and if
the sendNode meets the condition, it will install the metalink.
\par
For the other methods, such as \#onMessageSendIn: is sent to the breaker, it does
the same except for checking the condition. \#OnMessageSend installs breakpoints
in all methods, instead of the specified one.
\par
\#clear clears all breakpoints by going over the list of installed metalinks and
uninstalling them. This list is kept track of when sending the messages above to
the objectBreaker.
\par
Note that each of the methods mentioned above, besides \#clear checks if the
object has super sends. If it does it will throw an exception.
\section{Multiple exception handlers}
For this part of the project we are focussing on adding multiple exception
handlers to BlockClosure. Following the code snippet in the assignment
description, we add a method to BlockClosure to convert it to a MyBlockClosure.
\subsection{Requirement 1}
Once a BlockClosure is a MyBlockClosure, it unlocks methods relevant for this
part of the project. It unlocks methods such as message:with:, which allows the
user to register handlers for exception classes. It does this by keeping them as
an association in a collection internally. Initially, it tried to find the
closes matching recorded Exception class, and uses the associated handler.
\subsection{Requirement 2}
To call the handlers in order of specificity, we change the way that
MyBlockClosure finds the handler. First we create a SortedCollection from the
OrderedCollection with associations, sorted by hierarchy and filtered if the
thrown Exception is a subclass of it as well. After this, it simply throws the
first one (the one that is the most specific to the exception thrown).
\subsection{Requirement 3}
\label{subsection:requirement-3}
To prevent breaking of on:do:, we essentially need to allow MyBlockClosure to
allow an exception pass through if no handler can be found. This is done by
simply allowing the exception to pass, if no handler can be found.
\subsection{Requirement 4}
Finally, to allow calling the next handler in MyBlockClosure, we modify
MyBlockClosure, to keep track of the sorted collection of handlers as well as
the handler currently being used. When value: is sent, it sets these values and
when callNextHandler is sent, the index is incremented and the next handler in
the collection gets
used. If there is no handler left, it defaults to the behavior described in
\ref{subsection:requirement-3}
\end{document}
